Fast Regression of the Tritium Breeding Ratio in Tokamak Fusion Reactors

### 2020-05-29 paper/results.tex

The results shown in~\fref{fig:scaling} suggest that in terms of regression
performance the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training points. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based approaches
clearly outperform ANNs. This can be observed
particularly on sets of sizes up to~\num{6000}.

@@

Consistent with our expectation, the shortest training times were achieved by
instance-based learning methods (e.g. KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. Our conjecture is that this effect may be caused
by a constant multi-threading overhead that possibly dominates the training process
on relatively small training sets.

@@

\subsubsection{Model Comparison}
\todo{This section needs to be shortened.}

In the fourth experiment proposed
in~\sref{sec:experiment-methodology}, we exploit previously collected information
to produce surrogates with desirable properties for practical use. We
aim to create models that yield: (a)~the best regression performance regardless
of other features, (b)~acceptable performance with the shortest mean
prediction time, or (c)~acceptable performance with the smallest training set.
To this end, we trained 8~surrogates that are presented in~\fref{fig:reg-performance}
and~\tref{tbl:exp4-detailed-results}.

\begin{figure}[h]
	\centering
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model6}
	\end{minipage}\hfill%
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model7}
	\end{minipage}

@@

	\caption{Regression performance of models 1-4 (row 1, from the left) and 5-8
		(row 2) trained in experiment~4 (model comparison), viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}

Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (no.~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline measured during run~1
(see~\tref{tbl:sampling-runs} for details). Satisfying
goal~(b), the fastest model (no.~2, ANN) achieved~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy
goal~(c) we also trained a more simplified model (no.~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

\begin{table*}[t]
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\setlength\tabcolsep{2pt}
	{\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	}
	\caption{Results of experiment~4. Here, figures are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\SI{7.777049573054314}{\second}$
		measured in the MC TBR model during run~1 (see~\tref{tbl:sampling-runs}).
		The best-performing metrics are highlighted in bold.}
	\label{tbl:exp4-detailed-results}
\end{table*}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig5_sintoy}
	\caption{Sinusoidal toy TBR theory over two continuous parameters,
	$n=1$.}
	\label{fig:sintoy}
\end{figure}

@@

@@

An increase in initial samples with increment held constant had a strong impact
on final surrogate precision, an early confirmation of basic functionality. An
increase in MCMC candidate samples was seen to have a positive but very weak
effect on final surrogate precision, suggesting that the runtime of MCMC on each
iteration can be limited for increased efficiency. The most complex dynamics
arose with the adjustment of sample increment, shown in~\fref{fig:qassincr}. For
each tested initial sample quantity $N$, the optimal number of step samples was seen to be well-approximated by $\sqrt{N}$. The plotted error trends suggest that
incremental samples larger than this optimum give slower model improvement on both the training and evaluation sets, and a larger minimum error on the evaluation set. This performance distinction is predicted to be even more significant when trained on the expensive MC model, where the number of sample evaluations will serve as the primary bottleneck for computation time.
\begin{figure}[h!]
    \centering
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6a_qassincrsamp.pdf}
    \end{minipage}%
    \hfill%
    \begin{minipage}{0.5\textwidth}
        \centering
        \includegraphics[width=1.1\linewidth]{fig6b_qassincrtime.pdf}
    \end{minipage}
    \caption{QASS absolute training error over total sample quantity (left) and number of iterations (right). MAE represents surrogate error on the adaptively-sampled training/test set, and E\_MAE on the independent evaluation sets.}
    \label{fig:qassincr}
\end{figure}

@@

\begin{figure}[h]
\centering
\begin{minipage}{.48\textwidth}
  %\centering
  \includegraphics[width=1.1\linewidth]{fig7_qasssampling.pdf}
  %\caption{Absolute training error for QASS, baseline scheme, and mixed scheme.}
  %\label{fig:qasssampling}
\end{minipage}%
\hfill%
\begin{minipage}{.48\textwidth}
  %\centering
  \includegraphics[width=1.1\linewidth]{fig8_qasssampling100k.pdf}
  %\caption{Absolute training error for QASS and baseline scheme, with 100k initial samples.}
  %\label{fig:qasssampling100k}
\end{minipage}
\end{figure}

>>>>

The results shown in~\fref{fig:scaling} suggest that in terms of regression
performance the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as
more training points are introduced. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based approaches
clearly outperform ANNs. This can be observed
particularly on sets of sizes up to~\num{6000}.

@@

Consistent with our expectation, the shortest training times were achieved by
instance-based learning methods (KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. We suspect that this effect may be caused
by a constant multi-threading overhead that dominates the training process
on relatively small sets.

@@

\subsubsection{Model Comparison}
In the fourth experiment, we aim to create models that yield:
(a)~the best regression performance regardless
of other features, (b)~acceptable performance with the shortest mean
prediction time, or (c)~acceptable performance with the smallest training set.
To this end, we trained 8~surrogates that are presented in~\fref{fig:reg-performance}
and~\tref{tbl:exp4-detailed-results}.
Having selected ANNs, GBTs, ERTs, RBFs and SVMs based on the results of
experiments~2-3, we utilised the best-performing hyperparameters.
In pursuit of goal~(a), the best approximator (no.~1,
ANN) achieved~$R^2=\num{0.998}$ and mean prediction
time~$\overline{t}_{\text{pred.}}=\SI{1.124}{\micro\second}$. These correspond
to a standard error~$S=\num{0.013}$ and a relative speedup~$\omega=\num{6916416} \times$
with respect to the MC TBR evaluation baseline. Satisfying
goal~(b), the fastest model (no.~2, ANN) achieved~$R^2=\num{0.985}$,
$\overline{t}_{\text{pred.}}=\SI{0.898}{\micro\second}$, $S=\num{0.033}$
and~$\omega=\num{8659251} \times$.
While these surrogates
were trained on the entire available set of~\num{500000} datapoints, to satisfy
goal~(c) we also trained a more simplified model (no.~4, GBT)
that achieved~$R^2=\num{0.913}$,
$\overline{t}_{\text{pred.}}=\SI{6.125}{\micro\second}$, $S=\num{0.072}$ and $\omega=\num{1269777} \times$
with a set of size only~\num{10000}.

\begin{table*}
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\caption{\label{tbl:exp4-detailed-results}Results of experiment~4. Here, figures are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\SI{7.777049573054314}{\second}$
		measured in the MC TBR model during run~1 (see~\tref{tbl:sampling-runs}).
		The best-performing metrics are highlighted in bold.}
	\setlength\tabcolsep{4pt}
	\begin{indented}
	\item[]
		\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	\end{indented}
\end{table*}

\begin{figure}
	\centering
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model6}
	\end{minipage}\hfill%
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model7}
	\end{minipage}

@@

	\caption{Regression performance of models 1-8 (from left to right, top to
		bottom) trained in the model comparison experiment, viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{fig5_sintoy}
	\caption{Sinusoidal toy TBR theory over 2 continuous parameters,
	$n=1$.}
	\label{fig:sintoy}
\end{figure}

@@

An increase in initial samples with increment held constant had a strong impact
on final surrogate precision, an early confirmation of basic functionality. An
increase in MCMC candidate samples was seen to have a positive but very weak
effect on final surrogate precision, suggesting that the runtime of MCMC on each
iteration can be limited for increased efficiency. The most complex dynamics
arose with the adjustment of sample increment, shown in~\fref{fig:qassincr}. For
each tested initial sample quantity $N$, the optimal number of step samples was seen to be well-approximated by $\sqrt{N}$. The plotted error trends suggest that
incremental samples larger than this optimum give slower model improvement on
both the training and evaluation sets, and a larger minimum error on the
evaluation set. This performance distinction is predicted to be even more
significant when trained on the expensive MC model, where the number of sample
evaluations will serve as the primary bottleneck for computation time.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig6a_qassincrsamp.pdf}
	\includegraphics[width=\linewidth]{fig6b_qassincrtime.pdf}
    \caption{QASS absolute training error over total sample quantity (left) and number of iterations (right). MAE represents surrogate error on the adaptively-sampled training/test set, and E\_MAE on the independent evaluation sets.}
    \label{fig:qassincr}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig7_qasssampling.pdf}
	\caption{\label{fig:qasssampling}Absolute training error for QASS, baseline scheme, and mixed scheme.}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig8_qasssampling100k.pdf}
	\caption{\label{fig:qasssampling100k}Absolute training error for QASS and baseline scheme, with 100k initial samples.}
\end{figure}
